Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-03-11 bbbq_article.tex
Comments visible in doc

% GvdB:
% Add one short paragraph, 
% where you reiterate the most important findings:
% In this study, we show that ...(and than basically 
% the abstract in slightly different words and slightly 
% more detail). End with a statement of why it is 
% important (vaccin development based on TMHs?)...

@@

As a data source, multiple
NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used,
which are 'gene', to find the gene names of membrane proteins, 
'dbSNP' (\cite{sherry2001dbsnp}) for SNPs associated with those genes
and 'protein', to obtain the sequence of proteins that SNPs act upon.
The 'dbSNP' contains 650 million 
catalogued non-redundant humane variations (called RefSNPs,
\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).
% GvdB: Not really relevant. As these are sequencing artefacts in the DNA, these should be randomly distributed and not affect the conclusions.
% GvdB: In addition, you use the most frequent SNPs, which minimizes the chance of those sequencing artifacts.
% It is estimated that approximately ten percent (ranging from 2 to 36 percent) 
% are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}),
% yet if this has been corrected for (if possible) in the last decade is unknown.

@@

% RJCB: moved this from Introduction to Discussion.
%
% \paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}
% 
% It is important that MHC-I presents both the peptides of the 
% (healthy) cell, as well as possible pathogen-derived peptides when the
% cell is infected. As described above, MHC-I presents TMH-derived epitopes 
% from the human host more often. It is unknown if MHC-I has the same
% dedication to present epitopes that stem from TMHs derived from
% proteins produced by pathogen, either would the pathogen
% be a virus or 
% a bacterium.
% 
% \paragraph{MHC-II is expected to present TMHs}
% 
% If presentation of TMHs on MHC-I would bring an evolutionary advantage 
% in the recognition of pathogens by the immune system, 
% it would follow that this is equally important for MHC-II, 
% especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
% response (\cite{novy2007cd4}). 
% The mechanism to detect foreign TMHs by MHC-II would be unknown, 
% similar to the discovery that MHC-I presents TMHs.
% 
% \paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}
% 
% It is unknown if MHC-II, like MHC-I, 
% presents TMH-derived epitopes as often, either 
% in humans,
% bacteria or viruses.
% 
% \paragraph{Selection undetectable in whole proteome}
% 
% In general, on would hope that evolutionary selection results in
% an immune system that as most attentive for loci that are
% essential for a virus, as these will be most conserved.
% In SARS-CoV-2, for example, there is preliminary evidence that the strongest
% selection pressure is upon residues that changes its 
% virulence (\cite{velazquez2020positive}).
% These loci, however, only account for a small part of a pathogen's proteome.
% Additionally, these essential parts differ widely between pathogens.
% Because of this scarcity and variance in targets, 
% one can imagine that the human immune system 
% is not tailored to detect these sites, 
% as hinted by upon by the aforementioned influenza study.
% 
% \paragraph{Selection may be detectable in TMHs}
% 
% TMHs, on the other hand, also have their function constraints, 
% yet can occur multiple time a pathogen's proteome.
% One can safely assume a pathogen's proteome contains multiple TMHs.
% Therefore, it may be beneficial for the host
% if its immune system would be more attentive towards TMHs.
% And maybe this has already happened: MHC-I already detects hydrophobic
% peptides. This feature, however, may also be caused by selection
% to detect hydrophobic regions in the soluble proteins of pathogens.
% It is unknown, when focusing on TMHs only, if a signal of selection
% can be detected.
% 
% \paragraph{Use of protein data in phylogenetics}
% 
% When using DNA sequences, one can use a skewed rate
% in non- versus synonymous mutation
% to detect the signature of selection (\cite{murrell2015gene}).
% Using AA sequences, however, has its advantages,
% as its is closer to the actual phenotype
% selection acts upon: DNA may never be transcribed to RNA,
% or its RNA may never be translated (\cite{diz2012proteomics}).
% When using a proteome in phylogenetic research,
% we know that the majority of proteins are selected to just
% maintain their function most of the time, where
% is the time spans there is selection, only a few AAs
% can actually increase the 'fitness' of the
% protein (\cite{anisimova2009investigating}).
% There, when generalizing the dynamics of mostly purifying selection (to maintain
% a protein's function) and a short duration of positive selection,
% those genes that are selected cannot be detected (\cite{yang2000statistical}).

@@

% We concluded that the
% epitopes that MHC-I presents are [as/not as] likely 
% to be derived from TMH within either a human host and its bacterial pathogen.
% Because a bacterium does not infect a cell, thus its peptides
% will not be presented by MHC-I, this result is [unexpected/expected]
% 
% We aimed our evolutionary experiment at TMHs, because these can
% be predicted well from a protein structure,
% are common structures and are present in all pathogens. 
% We could have done the same experiment on beta-turn,
% as also these can be predicted well (\cite{petersen2010netturnp}),
% are common structures and are present in all pathogens.
% 
% The human immune system and human pathogen are in an evolutionary
% arms race: our immune systems is selected for the detection
% of pathogens, whereas pathogens are selected to avoid detection.
% From a pathogen's point of view, however, this struggle 
% is of only minor importance:
% in seasonal influenza, for example, the selection pressure
% exerted by the immune system was only limited (\cite{han2019individual}).

>>>>

[GVDB:
 Add one short paragraph, 
 where you reiterate the most important findings:
 In this study, we show that ...(and than basically 
 the abstract in slightly different words and slightly 
 more detail). End with a statement of why it is 
 important (vaccine development based on TMHs?)...
]

@@

As a data source, multiple
NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used,
which are 'gene', to find the gene names of membrane proteins, 
'dbSNP' (\cite{sherry2001dbsnp}) for SNPs associated with those genes
and 'protein', to obtain the sequence of proteins that SNPs act upon.
The 'dbSNP' contains 650 million 
catalogued non-redundant humane variations (called RefSNPs,
\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).

@@

% \paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}

[RJCB: this paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]

It is important that MHC-I presents both the peptides of the 
(healthy) cell, as well as possible pathogen-derived peptides when the
cell is infected. As described above, MHC-I presents TMH-derived epitopes 
from the human host more often. It is unknown if MHC-I has the same
dedication to present epitopes that stem from TMHs derived from
proteins produced by pathogen, either would the pathogen
be a virus or 
a bacterium.

% \paragraph{MHC-II is expected to present TMHs}

[RJCB: this paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]

If presentation of TMHs on MHC-I would bring an evolutionary advantage 
in the recognition of pathogens by the immune system, 
it would follow that this is equally important for MHC-II, 
especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
response (\cite{novy2007cd4}). 
The mechanism to detect foreign TMHs by MHC-II would be unknown, 
similar to the discovery that MHC-I presents TMHs.

% \paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}

[RJCB: this paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]

It is unknown if MHC-II, like MHC-I, 
presents TMH-derived epitopes as often, either 
in humans,
bacteria or viruses.

% \paragraph{Selection undetectable in whole proteome}

[RJCB: this paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]

In general, on would hope that evolutionary selection results in
an immune system that as most attentive for loci that are
essential for a virus, as these will be most conserved.
In SARS-CoV-2, for example, there is preliminary evidence that the strongest
selection pressure is upon residues that changes its 
virulence (\cite{velazquez2020positive}).
These loci, however, only account for a small part of a pathogen's proteome.
Additionally, these essential parts differ widely between pathogens.
Because of this scarcity and variance in targets, 
one can imagine that the human immune system 
is not tailored to detect these sites, 
 as hinted by upon by the aforementioned influenza study.

% \paragraph{Selection may be detectable in TMHs}

[RJCB: this paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]

TMHs, on the other hand, also have their function constraints, 
yet can occur multiple time a pathogen's proteome.
One can safely assume a pathogen's proteome contains multiple TMHs.
Therefore, it may be beneficial for the host
if its immune system would be more attentive towards TMHs.
And maybe this has already happened: MHC-I already detects hydrophobic
peptides. This feature, however, may also be caused by selection
to detect hydrophobic regions in the soluble proteins of pathogens.
It is unknown, when focusing on TMHs only, if a signal of selection
can be detected.

% \paragraph{Use of protein data in phylogenetics}

[RJCB: this paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]

When using DNA sequences, one can use a skewed rate
in non- versus synonymous mutation
to detect the signature of selection (\cite{murrell2015gene}).
Using AA sequences, however, has its advantages,
as its is closer to the actual phenotype
selection acts upon: DNA may never be transcribed to RNA,
or its RNA may never be translated (\cite{diz2012proteomics}).
When using a proteome in phylogenetic research,
we know that the majority of proteins are selected to just
maintain their function most of the time, where
is the time spans there is selection, only a few AAs
can actually increase the 'fitness' of the
protein (\cite{anisimova2009investigating}).
There, when generalizing the dynamics of mostly purifying selection (to maintain
a protein's function) and a short duration of positive selection,
those genes that are selected cannot be detected (\cite{yang2000statistical}).

[RJCB: that was the last paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]

@@

[RJCB: this paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]
We concluded that the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its bacterial pathogen.
Because a bacterium does not infect a cell, thus its peptides
will not be presented by MHC-I, this result is [unexpected/expected]

[RJCB: this paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]
We aimed our evolutionary experiment at TMHs, because these can
be predicted well from a protein structure,
are common structures and are present in all pathogens. 
We could have done the same experiment on beta-turn,
as also these can be predicted well (\cite{petersen2010netturnp}),
are common structures and are present in all pathogens.

[RJCB: this paragraph I ruthlessly moved
from introduction to here, as was suggested by GVDB]
The human immune system and human pathogen are in an evolutionary
arms race: our immune systems is selected for the detection
of pathogens, whereas pathogens are selected to avoid detection.
From a pathogen's point of view, however, this struggle 
is of only minor importance:
in seasonal influenza, for example, the selection pressure
exerted by the immune system was only limited (\cite{han2019individual}).
